\added{GrapeCPNet: A Self-supervised Point Cloud Completion Network for 3D Phenotyping of Grape Bunches}

### 2025-02-16 main.tex
discussion part 1

\added{Another challenge after the success of instance berry segmentation is the incomplete representation of berry surfaces.} 
The aggregated cluster growth causes the berries to squeezed against each other, leading to self occlusion and resulting in an incomplete point cloud of berry surfaces \citep{du_instance_2023}. 
Therefore, point cloud instance segmentation can only capture partial information rather than that of whole berries.
To address \replaced{this issue}{the issue of partial point clouds,} and consequently an underestimation of the berry traits, point cloud shape completion is one of the most commonly used solutions.
Point cloud completion is the task of generating a complete 3D shape given partial point cloud input \citep{tesema_point_2024}. 
For instance, \citet{ge_symmetrybased_2020} assume that strawberries are symmetrical, suggesting that their shape can be completed by mirroring the existing shape, provided an appropriate symmetrical plane can be identified.
\citet{marangoz_fruit_2022} proposed an super-ellipsoids matching approach to map fruits on plants and estimate their shape.
However, these methods are limited to specific fruit characteristics and require manual design of algorithms according to the features of the fruit, lacking the potential for generalized application.
Utilizing deep learning methods, researchers have designed point cloud completion networks \citep{wang_unsupervised_2021}. 
\citet{park_deepsdf_2019} introduced DeepSDF to achieve the completion of object point cloud models. It uses a learned continuous signed distance function to represent a class of shapes in latent space and is tolerant to noisy 3D input data. 
\citet{magistri_contrastive_2022} utilized DeepSDF and its latent space in an encoder-decoder network that converted a single RGB-D frame to a 3D shape of a complete fruit. 
This encoder-decoder network was extended and optimized by \citet{blok_highthroughput_2025} for high-throughput 3D shape completion of potato tubers on a commercial harvester. Outside the agricultural domain, \citet{tang_lakenet_2022} proposed a point cloud completion network (Lake-net) for furniture model point cloud datasets. 
For grape berry application, the shape of the berry can be roughly approached by the deformation of a sphere or ellipsoid depending on the species. 
Existing datasets from other fruits cannot be applied directly, and manually creating paired training datasets is time-consuming and labor-intensive.
Therefore, there is a demand to develop an analysis pipeline, capable of completing the detailed surface of an incomplete berry point cloud while requiring less labor-intensive training data preparation.

@@

\subsubsection{Berry instance segmentation}
\label{sec:251}
In this paper, we used average precision AP as the evaluation metric for berry instance segmentation, similar to the original paper of \added{SoftGroup} \citet{vu_softgroup_2022}. The average APs with IoU (Intersection over Union) ranging from 50\% to 95\% in steps of 5\% were used and noted as \replaced{mAP@[.5:.95]}{AP@50:5:95}.

@@

\subsection{Berry Instance Segmentation Results}

@@

% table 4
\begin{table}[h]
    \centering
    \caption{\replaced{Comparing proposed method with}{Experimental results of} different completion networks \deleted{based on self-supervised training} for berry point cloud. Best performing results are visualized in bold.}
    \begin{tabular}{cccccc}
        \hline
        \textbf{Model} & \textbf{DataSet} & \textbf{$d_{CD}$ $\downarrow$} & \textbf{Precision $\uparrow$} & \textbf{Recall $\uparrow$} & \textbf{F-score $\uparrow$} \\
        \hline
        \multirow{2}{*}{\makecell{PoinTr \\ \citep{yu_pointr_2021}}} & Test & 9.036 & 0.721 & 0.705 & 0.713 \\
        \cline{2-6}
        & Inference & 10.542 & 0.623 & 0.601 & 0.612 \\
        \hline
        \multirow{2}{*}{\makecell{SnowFlakeNet \\ \citep{xiang_snowflakenet_2021}}} & Test & 11.235 & 0.512 & 0.534 & 0.521 \\
        \cline{2-6}
        & Inference & 12.384 & 0.415 & 0.423 & 0.419 \\
        \hline
        \multirow{2}{*}{\makecell{TopNet \\ \citep{tchapmi_topnet_2019}}} & Test & 23.019 & 0.201 & 0.200 & 0.201 \\
        \cline{2-6}
        & Inference & 31.880 & 0.155 & 0.154 & 0.154 \\
        \hline
        \multirow{2}{*}{\makecell{GRNet \\ \citep{xie_grnet_2020}}} & Test & 19.240 & 0.402 & 0.392 & 0.397 \\
        \cline{2-6}
        & Inference & 21.250 & 0.365 & 0.363 & 0.364 \\
        \hline
        \multirow{2}{*}{\makecell{GrapeCPNet \\ (Ours)}} & Test & \textbf{8.315} & \textbf{0.729} & \textbf{0.772} & \textbf{0.751} \\
        \cline{2-6}
        & Inference & \textbf{9.158} & \textbf{0.685} & \textbf{0.696} & \textbf{0.689} \\
        \hline
    \end{tabular}
    \label{tbl:4}
\end{table}

@@

The radius parameters were calculated by fitting the berry to the surface of an ellipsoid\deleted{, as shown in Figure~\ref{fig:raw16}}.
From Figure~\ref{fig:raw17}, it can be seen that the proposed method achieves an average $R^2$ of 0.86 \added{and average MAE of 0.608 mm} for the three axis lengths of the four species of grape.

This result can be dissected as follows:

@@

(1) The different species: Red Grape, Green Grape, Kyoho Grape, and Shine-Muscat Grape axis length prediction reached an average $R^2$ of 0.86, 0.84, 0.89, and 0.84, respectively. 
Among them, Kyoho Grape achieved the best results due to its spherical berries that are not too densely clustered and the lower difficulty in berry completion and morphology fitting. 
Secondly, Red Grape and Green Grape both present a typical ellipsoid shape, but the \replaced{Red}{latter} is less precise than the \replaced{Green}{former} due to its \replaced{larger}{greater} ratio of long axis to short axis, which provides a more complex presentation. 
\added{As reported by \citet{schneider_predicting_2020}, the mean length-to-width ratio of Green Grape is 21.18mm to 18.75mm, while for Red Grape, it is 24.11mm to 18.41mm.}
Shine-Muscat Grape had the lowest precision, which was due to the fact that its berry grows tightly and occlude each other heavily, making it harder to accurately complete its shape.

@@

The experimental results showed that the proposed method achieved \added{both} high accuracy in the calculation of the radius \added{and volumes} of different grape species.
\added{In out study, we achieved an average $R^2$ of 0.86 and an average MAE of 0.608 mm across the three axis of berry, as well as an $R^2$ of 0.9689 and an MAE of 0.3842 $cm^3$ for berry volumes.}
\added{By comparision, \citet[Fig.~4]{rist_highprecision_2018} reported $R^2$ values of 0.54, 0.85, and 0.83 for berry width, length, and volume, respectively.}
\added{\citet{botturi_stewie_2023} reported a best-case MAE of 21.7mm for berry ratio estimataion using deep learning network on 2D images.}
\added{\citet{ni_threedimensional_2021} achieved a RMSE of 0.292 $cm^3$ for volume estimation.}
\added{Although the method by \citet{ni_threedimensional_2021} achieved a lower errors for volume estimataion, it relies on sphere fitting that are primarily applicable to clustered, round fruits.}
\added{In contrast, our proposed method not only accounts for occluded areas but is also adaptable to fruits of varying shapes, if appropriate paired training data is available.}

@@

% achievement and contribution of ths study
In this paper, \replaced{an automated deep learning pipeline for }{a method for the automatic }acquisition of 3D phenotypic traits of grape \added{bunch} was presented to address the problems of high labor cost, poor accuracy, and inability to realize end-to-end analysis. 
Based on the \added{3D} point cloud of a grape bunch \added{obtained by close-range multi-view photogrammetry}, a combination of deep learning point cloud processing algorithms realized the counting of the number of berries, the calculation of berry radius, and berry volume. 
\replaced{As a major contribution}{Moreover}, to address the problem of incomplete point clouds of berries due to occlusion, we introduced GrapeCPNet, a berry point cloud completion model based on the self-supervised training method.

% revisiting background to reboot reader

% revisit gap/aim/method

% revisit results and explore their implecations

% amap to literatures for comparison/support

% ppotential limitations and suggestions for future works

% the achievement / impact of study / potential applications.
As discussion point, we found that GrapeCPNet was not effective in the case of complex berry appearance, such as drop-shaped. 
We have found that some researchers \citep{pan_panoptic_2023,magistri_efficient_2024} have improved the completion effect by encoding a prior knowledge of the geometry of the object into the weights of the neural network. 
We subsequently consider utilizing the prior knowledge of berry geometry to achieve more detailed morphological feature extraction and representation to improve the quality of completion. 
In addition, the noise in the data is also related to the effectiveness of the method proposed in this paper. 
How to implement the method in practical application scenarios, such as RGB-D collected point cloud data, is another issue to be considered in future research.

@@

\section{Conclusions}

\added{In this study, we proposed a deep-learning-based pipeline for processing 3D phenotyping grape bunches reconstructed by photogrammetry.}
\added{The major contribution is the self-supervised completion network, named GrapeCPNet, designed to address the problem of missing parts in berries occluded within grape bunches with few manual annotations.}
\added{Due to the dense and compact arrangement of berries within a grape bunch, the inner portions are often occluded and not visible.}
\added{This often presents challenges in algorithm design for individual berry segmentation and accurate berry size estimation.}
\added{We utilized state-of-the-art, data-driven deep learning approaches to address these challenges.}
\added{First, the SoftGroup network, utilizing segmentation annotations from 8 grape bunches, was used to train a data-driven berry instance segmentation model.}
% \added{By randomly producing missing areas from complete berries, we automatically generated large number of training data for completion tasks.}
\added{The proposed GrapeCPNet also achieved better performance than existing completion networks, with Precision of 0.729, Recall of 0.772 and F-score of 0.751.}
\added{The ellipsoid surface fitting, used to calculate three axis lengths and volumes, achieved an average $R^2$ of 0.86 and 0.97 respectively, when comparing completed berries with their ground truth.}
\added{This study clearly demonstrates the potential of using data-drive strategies with minimal annotation effort to address the 3D occlsuion challenge in agricultural applications.}
\added{Although this network has only been tested on grape bunches collected in a controlled indoor environment, the proposed deep learning pipelines and training data generation approach have broad potential applications for in-field vineyards and even other crops in the future.}

>>>>

\added{Another challenge after the success of instance berry segmentation is the incomplete representation of berry surfaces.} 
The aggregated cluster growth causes the berries to squeezed against each other, leading to self occlusion and resulting in an incomplete point cloud of berry surfaces \citep{du_instance_2023}. 
Therefore, point cloud instance segmentation can only capture partial information rather than that of whole berries.
To address \replaced{this issue}{the issue of partial point clouds,} and consequently an underestimation of the berry traits, point cloud shape completion is one of the most commonly used solutions.
Point cloud completion is the task of generating a complete 3D shape given partial point cloud input \citep{tesema_point_2024}. 
For instance, \citet{ge_symmetrybased_2020} assume that strawberries are symmetrical, suggesting that their shape can be completed by mirroring the existing shape, provided an appropriate symmetrical plane can be identified.
\citet{marangoz_fruit_2022} proposed an super-ellipsoids matching approach to map fruits on plants and estimate their shape.
However, these methods are limited to specific fruit characteristics and require manual design of algorithms according to the features of the fruit, lacking the potential for generalized application.
Utilizing deep learning methods, researchers have designed point cloud completion networks \citep{wang_unsupervised_2021}. 
\citet{park_deepsdf_2019} introduced DeepSDF to achieve the completion of object point cloud models. It uses a learned continuous signed distance function to represent a class of shapes in latent space and is tolerant to noisy 3D input data. 
\citet{magistri_contrastive_2022} utilized DeepSDF and its latent space in an encoder-decoder network that converted a single RGB-D frame to a 3D shape of a complete fruit. 
This encoder-decoder network was extended and optimized by \citet{blok_highthroughput_2025} for high-throughput 3D shape completion of potato tubers on a commercial harvester. 
\deleted{Outside the agricultural domain, tang\_lakenet\_2022 proposed a point cloud completion network (Lake-net) for furniture model point cloud datasets.} 
For grape berry application, the shape of the berry can be roughly approached by the deformation of a sphere or ellipsoid depending on the species. 
Existing datasets from other fruits cannot be applied directly, and manually creating paired training datasets is time-consuming and labor-intensive.
Therefore, there is a demand to develop an analysis pipeline, capable of completing the detailed surface of an incomplete berry point cloud while requiring less labor-intensive training data preparation.

@@

\subsubsection{Berry instance segmentation}
\label{sec:251}
In this paper, we used average precision AP as the evaluation metric for berry instance segmentation, similar to the original paper of \added{SoftGroup} \citep{vu_softgroup_2022}. The average APs with IoU (Intersection over Union) ranging from 50\% to 95\% in steps of 5\% were used and noted as \replaced{mAP@[.5:.95]}{AP@50:5:95}.

@@

\subsection{Berry Instance Segmentation Results}
\label{sec: insegresults}

@@

% table 4
\begin{table}[h]
    \centering
    \caption{\replaced{Comparing proposed method with}{Experimental results of} different completion networks \deleted{based on self-supervised training} for berry point cloud. Best performing results are visualized in bold.}
    \begin{tabular}{cccccc}
        \hline
        \textbf{Model} & \textbf{DataSet} & \textbf{$d_{CD}$ (mm) $\downarrow$} & \textbf{Precision $\uparrow$} & \textbf{Recall $\uparrow$} & \textbf{F-score $\uparrow$} \\
        \hline
        \multirow{2}{*}{\makecell{PoinTr \\ \citep{yu_pointr_2021}}} & Test & 9.036 & 0.721 & 0.705 & 0.713 \\
        \cline{2-6}
        & Inference & 10.542 & 0.623 & 0.601 & 0.612 \\
        \hline
        \multirow{2}{*}{\makecell{SnowFlakeNet \\ \citep{xiang_snowflakenet_2021}}} & Test & 11.235 & 0.512 & 0.534 & 0.521 \\
        \cline{2-6}
        & Inference & 12.384 & 0.415 & 0.423 & 0.419 \\
        \hline
        \multirow{2}{*}{\makecell{TopNet \\ \citep{tchapmi_topnet_2019}}} & Test & 23.019 & 0.201 & 0.200 & 0.201 \\
        \cline{2-6}
        & Inference & 31.880 & 0.155 & 0.154 & 0.154 \\
        \hline
        \multirow{2}{*}{\makecell{GRNet \\ \citep{xie_grnet_2020}}} & Test & 19.240 & 0.402 & 0.392 & 0.397 \\
        \cline{2-6}
        & Inference & 21.250 & 0.365 & 0.363 & 0.364 \\
        \hline
        \multirow{2}{*}{\makecell{GrapeCPNet \\ (Ours)}} & Test & \textbf{8.315} & \textbf{0.729} & \textbf{0.772} & \textbf{0.751} \\
        \cline{2-6}
        & Inference & \textbf{9.158} & \textbf{0.685} & \textbf{0.696} & \textbf{0.689} \\
        \hline
    \end{tabular}
    \label{tbl:4}
\end{table}

@@

The radius parameters were calculated by fitting the berry to the surface of an ellipsoid\deleted{, as shown in Figure~\ref{fig:raw16}}.
From Figure~\ref{fig:raw17}, it can be seen that the proposed method achieves an average $R^2$ of 0.86 \added{and average MAE of 0.608 mm} for the three axis lengths of the four species of grape. 
This result can be dissected as follows:

@@

(1) The different species: Red Grape, Green Grape, Kyoho Grape, and Shine-Muscat Grape axis length prediction reached an average $R^2$ of 0.86, 0.84, 0.89, and 0.84, respectively. 
Among them, Kyoho Grape achieved the best results due to its spherical berries that are not too densely clustered and the lower difficulty in berry completion and morphology fitting. 
Secondly, Red Grape and Green Grape both present a typical ellipsoid shape, but the \replaced{Red}{latter} is less precise than the \replaced{Green}{former} due to its \replaced{larger}{greater} ratio of long axis to short axis, which provides a more complex presentation. 
\added{As also reported by \citet{schneider_predicting_2020}, the mean length-to-width ratio of Green Grape is 21.18mm to 18.75mm, while for Red Grape, it is 24.11mm to 18.41mm.}
Shine-Muscat Grape had the lowest precision, which was due to the fact that its berry grows tightly and occlude each other heavily, making it harder to accurately complete its shape.

@@

The experimental results showed that the proposed method achieved \added{both} high accuracy in the calculation of the radius \added{and volumes} of different grape species.
\added{In out study, we achieved an average $R^2$ of 0.86 and an average MAE of 0.608 mm across the three axis of berry, as well as an $R^2$ of 0.9689 and an MAE of 0.3842 $cm^3$ for berry volumes.}
\added{By comparision, \citet[Fig.~4]{rist_highprecision_2018} reported $R^2$ values of 0.54, 0.85, and 0.83 for berry width, length, and volume, respectively.}
\added{\citet{botturi_stewie_2023} reported a best-case MAE of 21.7mm for berry ratio estimataion using deep learning network on 2D images.}
\added{Although the method by \citet{ni_threedimensional_2021} achieved a lower errors for volume estimataion (RMSE of 0.292 $cm^3$), it relies on sphere fitting that are primarily applicable to clustered, round fruits.}
\added{In contrast, our proposed method not only accounts for occluded areas but is also adaptable to fruits of varying shapes, if appropriate paired training data is available.}

@@

% achievement and contribution of ths study
In this paper, \replaced{an automated deep learning pipeline for}{a method for the automatic} acquisition of 3D phenotypic traits of grape \added{bunch} was presented to address the problems of high labor cost, poor accuracy, and inability to realize end-to-end analysis. 
\replaced{As a major contribution}{Moreover}, to address the problem of incomplete point clouds of berries due to occlusion, we introduced GrapeCPNet, a berry point cloud completion model based on the self-supervised training method.
\deleted{Based on the point cloud of a grape bunch, a combination of deep learning point cloud processing algorithms realized the counting of the number of berries, the calculation of berry radius, and berry volume. }

% revisiting background to reboot reader
\added{Grape berry counting and size estimation plays a critical role in vineyard management and yield prediction.}
\added{By using common digital cameras alongside deep learning networks, several studies have achieved automatic berry detection and instance segmentation \citep{jadhav_comparative_2021, chen_instance_2023}, and then forward the analyses of geometry traits and yield estimation \citep{quinones_gcnet_2025}.}
\added{However, because of the dense and compact arrangement of berries within a grape bunch, some portions are often occluded and not visible when relying solely on single 2D images.}
\added{To reduce the effects of occlusion, \citet{du_instance_2023} reported accurate berry counting estimations by correlating image-based detections with ground measurements.}
\added{However, applying this method to other vineyards requires extensive, labor-intensive ground measurements}
\added{Similarly, \citet{botturi_stewie_2023} attempted to extract feature maps, such as berry radii, directly from deep networks applied to 2D images}
\added{Despite this approach, it still suffers from information loss and resulted in a high MAE of 21.7 mm.}
\added{Applying multi-view images for 3D reconstruction can significantly reduce the extent of occlusion caused by viewpoint limitations.}
\added{But unlike 2D images, which have well-structured data in the form of organized pixel matrices, the unordered nature of points in a 3D point cloud introduces significant challenges.}
\added{Meanwhile, the inner portion of berry is still invisible, only the surface of exposed berry surfaces can be scanned or reconstructed.}
\added{This complicates the algorithm design for individual berry segmentation and accurate berry size estimation.}

% revisit gap/aim/method
\added{For individual berry segmentation, early studies relied on the circular characteristics of the visible portions of berries for segmentation.}
\added{Building on the work of \citet{schnabel_efficient_2007}, who introduced point cloud shape detection using RANSAC, researchers were able to cluster partial berries based on their rounded shapes \citep{scholer_automated_2015, rist_highprecision_2018, schneider_predicting_2020}.}
\added{However, this feature-based approach performs well only in sparse grape arrangements and proves ineffective in more complex, dense conditions.}
\added{The reliance on manually designed features for algorithm development has now reached its limitations.}
\added{With advancements in machine learning, the focus has shifted to a data-driven era, where algorithms can learn complex feature patterns automatically.}
\added{For instance, \citet{rose_automated_2016} applied machine learning to distinguish features between grape berry clusters and background leaves automatically.}
\added{Similarly, \citet{ni_threedimensional_2021} implemented a two-step approach by first using mature 2D instance segmentation techniques on 2D images and then projecting the resulting segmentation masks onto reconstructed 3D models to complete the challenging task of 3D instance segmentation.}
\added{In another approach, \citet{luo_3d_2023} Mask R-CNN for 2D image segmentation and projected the results onto RGB-D point clouds.}
\added{The normal directions derived from the point clouds were then packed as an additional feature in the PointNet++ network for point cloud segmentation.}
\added{In this study, we applied the state-of-the-art point cloud segmentation network, SoftGroup \citep{vu_softgroup_2022} directly to the problem.}
\added{The primary time-consuming work of our approach was the annotation of 3D data for instance segmentation.}
\added{Compared to similar studies on grape berries, our proposed method demonstrated superior performance, achieving the best results among the reviewed methods (Subsection~\ref{sec: insegresults}).}

\added{For berry completetion by deep learning, the most difficult and time-consuming work is preparing the paired training dataset with complete and incomplete point cloud.}
\added{As our previous studies conducted for potato tuber completion networks \citep{blok_highthroughput_2025}, a distinguishable pin was used as reference point for pairing complete tubers by photogrammetry and incomplete tubers by RGBD camera.}
\added{Then the pin-guided point cloud matching algorithm is applied to calculate transformation matrix between paired dataset.}
\added{The potato tuber is single object with clear texture and shape variances, thus pairing through two different scans by two sensors is not too difficult.}
\added{For the grape berry, it is super rounded with weak texture features and strong reflectance, making the paired data super challenging.}
\added{Thus, the previous pin-guided point cloud paring methods for collection paired data does not work.}

% the achievement / impact of study / potential applications.
\deleted{As discussion point, we found that GrapeCPNet was not effective in the case of complex berry appearance, such as drop-shaped.}

% future works.
We have found that some researchers \citep{pan_panoptic_2023,magistri_efficient_2024} have improved the completion effect by encoding a prior knowledge of the geometry of the object into the weights of the neural network. 
We subsequently consider utilizing the prior knowledge of berry geometry to achieve more detailed morphological feature extraction and representation to improve the quality of completion. 
In addition, the noise in the data is also related to the effectiveness of the method proposed in this paper. 
How to implement the method in practical application scenarios, such as RGB-D collected point cloud data, is another issue to be considered in future research. \citet{blok_highthroughput_2025}

@@

\section{Conclusion}

\added{In this study, we proposed a deep-learning-based pipeline for processing 3D phenotyping grape bunches reconstructed by photogrammetry.}
\added{The major contribution is the self-supervised completion network, named GrapeCPNet, designed to address the problem of missing parts in berries occluded within grape bunches with few manual annotations.}
\added{We utilized state-of-the-art, data-driven deep learning approaches to address these challenges.}
\added{First, the SoftGroup network, utilizing segmentation annotations from 8 grape bunches, was used to train a data-driven berry instance segmentation model.}
% \added{By randomly producing missing areas from complete berries, we automatically generated large number of training data for completion tasks.}
\added{The proposed GrapeCPNet also achieved better performance than existing completion networks, with Precision of 0.729, Recall of 0.772 and F-score of 0.751.}
\added{The ellipsoid surface fitting, used to calculate three axis lengths and volumes, achieved an average $R^2$ of 0.86 and 0.97 respectively, when comparing completed berries with their ground truth.}
\added{This study clearly demonstrates the potential of using data-drive strategies with minimal annotation effort to address the 3D occlsuion challenge in agricultural applications.}
\added{Although this network has only been tested on grape bunches collected in a controlled indoor environment, the proposed deep learning pipelines and training data generation approach have broad potential applications for in-field vineyards and even other crops in the future.}
